---
title: "Session B - Identifying distributions."
author: <INSERT NAME>
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Throughout this course we will use a package called 'tidyverse'. This is a 'superpackage' which contains lots of other packages with lots of functions. If you haven't installed this already then you need to go ahead and run the line below. You only need to install a package once and afterwards should use a '#' before it to stop it from ever running again.

```{r Tidyverse Install}
#install.packages("tidyverse")
#install.packages("tidyverse") <- Make it look like this once you have it installed.
```

You will notice in your console that the section above appeared in a 'grey' area. This is because this is an R Markdown document. This is represented by the file extension .Rmd. Meanwhile there are also R Scripts (represented by .R extensions). Typically when writing code for publishable purposes or for software we have a series of R Scripts, however R markdown files are becomming more flavoursome when examining unqiue datasets since a neater overview can be given, it also allows for a neat PDF (or other document type) to be an output showing text, code and code output.

Having installed tidyverse earlier, we must still load in the package for our system to use it. We do that with the 'library' command - this is a built in R function. Let's do that now below, noting we do not need to use the quotes to load it in:

```{r Tidyverse Loading}
library(tidyverse)
```

Now that we know how to summarise individual variables, we may want to start to look at relationships between variables. The go-to method to do this is a scatterplot. Run the code below and see what it does.

```{r ScatterPlot}
plot1 <- cats %>%
  ggplot(aes(x = Bwt, y = Hwt)) +
  geom_point()

plot1
```

### Exercise 4
Comment on the relationship between body weight and heart weight, using the plot.

Your Answer:

-------------

### Exercise 5
Amend the code given above, to colour the points on the plot by the sex of the cat. It may be useful to look at the documentation for the ggplot package. We do this by placing a question mark next to the package name and running the code, ie. "?ggplot".

Your Answer:
```{r ScatterPlot2}

```
-------------

### Exercise 6
Once again amend the code. This time make it so that there are two plots which show two different the points seperated by the sex of the cats. Do you notice anything different from when they were on one plot? Write a short explanation.
[Hint: Look up what 'facet_wrap' does.]

Your Answer:
```{r ScatterPlot3}

```

-------------