---
title: "Session B - Identifying distributions. (SOLUTIONS)"
author: <INSERT NAME>
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Throughout this course we will use a package called tidyverse, a 'superpackage' containing tools for data manipulation and visualisation. Remember you only need to install a package once, and afterwards you should comment out the install line.

```{r Tidyverse Install}
# install.packages("tidyverse")
# install.packages("palmerpenguins")
```

```{r Tidyverse Loading}
library(tidyverse)
library(palmerpenguins)
```

In Session A we summarised variables using means, medians, modes and standard deviations. Now we move to identifying the shape of a distribution, as introduced in the slides (e.g. symmetric, skew left, skew right, multimodal)

A histogram is the simplest tool for this.

```{r HistogramBasic}
ggplot(penguins, aes(x = bill_length_mm)) +
  geom_histogram(bins = 20, fill = "steelblue", colour = "white")
```

### Exercise 1
Looking at the histogram above, describe the distribution of bill length. Is it symmetric, skewed, or multimodal?

Your Answer:

-------------

Density plots give a smooth approximation of the histogram.

```{r DensityPlot}
ggplot(penguins, aes(x = bill_length_mm)) +
  geom_density(fill = "orange", alpha = 0.4)
```

### Exercise 2
Compare the histogram and the density plot. What additional detail does the density plot reveal, if any?

Your Answer:

-------------

### Exercise 3
We want to see how different species vary in bill length. Create three separate histograms, one for each species.
Use facet_wrap(~species) to help you.

Your Answer:
```{r GroupedHistExercise}

```

-------------

### Exercise 4

Create your own scatterplot of:
- x-axis: flipper_length_mm
- y-axis: bill_length_mm

Then add transparency (alpha = 0.6) and increase point size.

Your Answer:
```{r ScatterplotExercise}

```

-------------

### Exercise 5
A Q-Q plot helps identify whether a variable follows a normal distribution. Create a Q-Q plot for bill_length_mm. 

Your Answer:
```{r QQPlotExercise}

```

-------------

Excellent — you've now actively coded multiple tools for identifying and comparing distributions.
